---
title: "R programming"
output: html_document
date: '2022-06-18'
---

# Semana 1

```{r}
getwd()
```

```{r}
dir()
```

```{r}
monda <- matrix(1, nrow = 3, ncol= 3)
print(monda)
```

The \<- symbol is the assignment operator

```{r}
mensaje_mpva <- 'ojalá algun día nos casemos'
mensaje_mpva # This is called auto-printing 
print(mensaje_mpva) # This is explicit printing 

```

```{r}
x <- 1:20
x
```

Los numeros entre [] indican la posición del primer objeto a la derecha

# Data Types

R has five basic classes of objects:

-   character
-   numeric (real numbers)
-   integer
-   complex
-   logical (True/False)

The most basic object is a vector: contains multiple copies of a single type of objects. So you cannot mix different objects under the same vector

BUT: the one exception is a *list*, which is represented as a vector put can contain objects of different classes

Empty vectors can be created:

```{r}
vector()
```

## Numbers

Numbers in R are generally treated a numeric objects (i.e. double precision real numbers)

If you explicitly want an integer, you need to specify the L suffix

```{r}
num_entero <- 1L 
typeof(num_entero)
num <- 1 
typeof(num)
```

## Attributes

R objects can have attributes

-   names, dimnames
-   dimensions
-   class
-   length

Attributes of an object can be accessed using the attributes( ) function

```{r}
attributes(monda)
```

## Vectors and Lists

Creating vectors

The c() function can be used to create vectors of objects

```{r}
x <- c(0.5, 0.6) ##numeric
x <- c(T,F) ##logical
x <- c('a', 'b', 'c') ##character 
x <- 9:29 ##numeric or integer?
typeof(x)
x <- c(1+0i, 2+4i) ##complex 

```

Si sería un integer

```{r}
x <- vector('numeric', length=10)
x
```

When different objects are mixed in a vector, coercion occurs so that every element in the vector is of the same class

```{r}
y <- c(1.7, 'a')
y
```

En este caso el número 1.7 se transformó en una string

### Explicit Coercion

Objects can be explicitly coerced fron one class to another usint the as.\* function. For example:

```{r}
x <- 0:7
typeof(x)
as.numeric(x)
as.logical(x)
as.character(x)


```

## Lists

Lists are a special type of vector that can contain elements of different classes.

```{r}
x <- list('a', F, 2.65)
x
```

## Matrices

Matrices are vectors with a dimmension attribute

```{r}
m <- matrix(1:20,nrow = 4, ncol=5)
dim(m)
m
```

Las matrices se llenan por columnas. Es decir los valores se asignan a la primera columna hasta que alcanza su dimensión y luego empieza a rellenar los valoresde la segunda columna y así.

También podría añadirle la dimensión al objeto y luego convertirlo en una matriz

```{r}
m <- 1:20
m
dim(m) <- c(4,5)
m
```

column binding

```{r}
x <- 1:4
y <-11:14

cbind(x,y)

```

row binding

```{r}
x <- 2:4
y <- 12:14
rbind(x,y)
```

## Factors

Factors are used to represent categorical data. Factors can be unordered or ordered. One can think of a factor as an integer vector where each integer has a label.

Using factors with labels is better than using integers because factors are self-describing; having a variable that has values 'Male' and 'Female' is better tah having a variable that has values 1 and 2.

```{r}
x <- factor(c('yes', 'yes', 'no', 'no', 'yes'))
x
table(x)
unclass(x)
```

The order of the levels can be set using the levels argument to facotr().

```{r}
x <- factor(c('yes', 'yes', 'no', 'no', 'yes'), levels = c('yes', 'no'))
x
table(x)
unclass(x)

```

Lo que hice acá fue cambiar el valor categórico que tiene las observaciones. yes es 1 en vez de ser 2.

## Msiing Values

Missing values are denoted by NA or NaN for undefined mathmatical operations

is.na() is used to test objects if they are NA

is.nan() is ued to test for NaN

NA values have a class also.

A NaN value is also NA but the converse is not true

```{r}
x <- c(1,2,NA,10,3)
is.na(x)
is.nan(x)

```

```{r}
x<- c(1,2, NaN, NA, 4)
is.na(x)
is.nan(x)

```

## Data Frames

Special type of list where very element of the list has to have the same length

They can also store different types of objects

```{r}
data.matrix(m)

```

```{r}
d <- data.frame(foo = 1:4, bar = c(T,T,F,F))
d
```

## Reading Data

There are a few principal functions reading data into R

-   `read.table`, `read.csv`, for reading tabular data

-   `readLines` for reading lines of a text file

-   `source`, for reading R in code files (inverse of dump)

-   `dget`, for reading in R code files

-   `load` for reading in saved workspaces

-   `unserialize` for reading single R objects in binary form

### read.table

The `read.table` function is one of the most commonly used functions for reading data. It has a few important arguments:

-   file; the name of a file or a connection

-   header; logical indicating if the file has a header line

-   sep; a string indicating how the colums are separating

-   colClasses; a character vector indicating the class of each column in the dataset

-   nrows; the number of rows in the dataset

-   comment,char; a character string indicating the comment character

-   skip; the numebr of lines to skip from the beginning

-   stringAsFactors; should character be coded as factors?

### Reading Larger Datasets with read.table

## Textual Formats

- `dumping` and `dputing` are useful because the resulting textual format is edit-able and in the case of corruption, potenntially recoverable 

### d-put

dput function reconstructs an r object 


```{r}
df <- data.frame(a= 1, b='a')
dput(df)
```
En el siguiente chunk lo mandaré a un archivo y después lo voy a leer

```{r}
dput(df, file= 'y.R')
new.df <- dget('y.R')
new.df
```

### dump

multiple objecs can be deparsed usint the dump fucntiona nd read back in using `source`

```{r}
x <- 'foo'
y <- data.frame(a= 1, b= 'a')
dump(c('x','y'), file = 'data.R')
rm(x,y)
source('data.R')
y
x
```

## Connections: Interfaces to the outside world 

Data are read in suing ´*connection* interfaces. Connections can be made to files (most common) or to more exotic things 

- `file`; opens a connection to a file
- `gzfile`; opnes a connection to a file compressed eith gzip 
- `url`; opens a connection to a webpage 


## Subsetting 

There a number of operators that can be used to extract subsets of R objects 

- [ always returns an object of the same class as the original; cna be used to select more than one element 

- [[ is used to extract elements of a list or a data frame; it can only be used to extract a single element and the class of the returned object will not necessarily be a list or a data frame 

- $ is used to extract elements of a list or data frame by name; semantics are similar to that of [[

Example 



```{r}

x <- c('a', 'b', 'c', 'd', 'd', 'd', 'c')
x[1]
x[2]
x[1:4]
```

```{r}
# logical subsetting 
x[x>'a']
# create a logical vector by subsetting 
u <- x>'a'
u
```

### Subsetting - lists 

```{r}
x <- list(foo= 1:4, bar = 0.6)
x[1]

x[[1]]

x$bar
# estoy utilizando el nombre para extraer el elemento 

x[['bar']]
x['bar']
```

```{r}
x <- list(foo= 1:4, novia= 'mpva', cancion = 'locos', fecha= 11)

i <-x[c(1:4)]
i$novia
```
```{r}

```

### Subsetting Nested Elements of a list 

The [[ can take an integer sequence 


```{r}
x <- list( a= list(1,2,3,4), b= list(7,6,5,4), c= list(0,9,8,7))

# Existen dos formas de extraer elementos que se están anidados dentro de mpas de un objeto 

# La primera forma es:
x[[2]][[3]]

# La segunda forma es 

x[[c(3,4)]]
```

### Subsetting - Matrices 

```{r}
matriz <- matrix(1:8, 4,2)
matriz
# El siguiente comando extrae el elemento ubicado en la tercera fila y en la segunda columna 
matriz[3,2]

# El siguiente comando extrae los primeros 6 elementos de la matriz. Recordar que estas se enumeran por columna 
matriz[1:6]

matriz[1,] #primera fila
matriz[,2] #segunda columna 
```
By default, when a single element of a matrix is retrieved, it is returned as a vector of length 1 rather than a 1x1 matrix. This behavior can be turned off by setting drop =FALSE 

```{r}
matriz[2,2, drop = FALSE]
```

```{r}
matriz[,2, drop = F]
```

### Subsetting - partial matching 

Partial matching of names is allowed with [[ and $

El $ hace partial matching del elemento que más se parezca 

```{r}
x <- list(aafuera = 1:6)
x$a

```
```{r}
X[['a']]
```

```{r}
x[['a', exact = FALSE]]
```

### Subsetting - removing missing values

```{r}
x <- c(1,2, NA, 4, NA, 5, 'k', 'p')
bad <- is.na(x)
bad
x[!bad]
```

```{r}
y <- c('b', 'c', 1, 6, 'f', 'NA', NA, NA)

good <- complete.cases(x,y)

x[good]
y[good]
```

```{r}
airquality[1:6,1:4]

good <- complete.cases(airquality)
airquality[good, ][1:6,1:4]
```

## Vectorized operations 

Many operations in R are vectorized making code more efficient, concise, and easier to read 
```{r}
x <- 1:4; y <- 7:8

x + y
```

```{r}
x <- 4
class(x)
x <-c(4, 'a', T)
class(x)
x<- c(1,3,5)
y <-c (3,2,10)
x <- list(2, 'a', 'b', T)
class(x[[1]])
x <- c(76:80)
u <- x[x==4]>78
u
```

```{r}
g <- read.csv('hw1_data.csv')
g
nrow(g)
tail(g, 2)
g[47, 1]

ozone <- g[,1]
ozone

good <- complete.cases(ozone)
ozgood <- ozone[good]
ozgood <- list(ozgood)

sapply(ozgood, mean)
ozgood


df <- as.data.frame(g)
df2 <- subset(df, Ozone >31 & Temp > 90)
m <- mean(df2$Solar.R)
m<- subset(df,Month ==5)
m
mm <- max(m$Ozone)
mm
```


