---
title: "R programming"
output: html_document
date: '2022-06-18'
---

# Semana 1 

```{r}
getwd()
```

```{r}
dir()
```

```{r}
monda <- matrix(1, nrow = 3, ncol= 3)
print(monda)
```

The <- symbol is the assignment operator 

```{r}
mensaje_mpva <- 'ojalá algun día nos casemos'
mensaje_mpva # This is called auto-printing 
print(mensaje_mpva) # This is explicit printing 

```

```{r}
x <- 1:20
x
```

Los numeros entre [] indican la posición del primer objeto a la derecha 

# Data Types 

R has five basic classes of objects: 

- character 
- numeric (real numbers)
- integer 
- complex 
- logical (True/False)

The most basic object is a vector: contains multiple copies of a single type of objects. So you cannot mix different objects under the same vector 

BUT: the one exception is a *list*, which is represented as a vector put can contain objects of different classes 

Empty vectors can be created:

```{r}
vector()
```

## Numbers 

Numbers in R are generally treated a numeric objects (i.e. double precision real numbers)

If you explicitly want an integer, you need to specify the L suffix 

```{r}
num_entero <- 1L 
typeof(num_entero)
num <- 1 
typeof(num)
```

## Attributes 

R objects can have attributes

- names, dimnames
- dimensions 
- class
- length 


Attributes of an object can be accessed using the attributes( ) function 

```{r}
attributes(monda)
```

## Vectors and Lists 

Creating vectors 

The c() function can be used to create vectors of objects 

```{r}
x <- c(0.5, 0.6) ##numeric
x <- c(T,F) ##logical
x <- c('a', 'b', 'c') ##character 
x <- 9:29 ##numeric or integer?
typeof(x)
x <- c(1+0i, 2+4i) ##complex 

```

Si sería un integer 

```{r}
x <- vector('numeric', length=10)
x
```

When different objects are mixed in a vector, coercion occurs so that every element in the vector is of the same class 

```{r}
y <- c(1.7, 'a')
y
```

En este caso el número 1.7 se transformó en una string 

### Explicit Coercion 

Objects can be explicitly coerced fron one class to another usint the as.* function. For example:

```{r}
x <- 0:7
typeof(x)
as.numeric(x)
as.logical(x)
as.character(x)


```

## Lists 

Lists are a special type of vector that can contain elements of different classes.

```{r}
x <- list('a', F, 2.65)
x
```

## Matrices 

Matrices are vectors with a dimmension attribute 

```{r}
m <- matrix(1:20,nrow = 4, ncol=5)
dim(m)
m
```

Las matrices se llenan por columnas. Es decir los valores se asignan a la primera columna hasta que alcanza su dimensión y luego empieza a rellenar los valoresde la segunda columna y así. 

También podría añadirle la dimensión al objeto y luego convertirlo en una matriz

```{r}
m <- 1:20
m
dim(m) <- c(4,5)
m
```

column binding 

```{r}
x <- 1:4
y <-11:14

cbind(x,y)

```

row binding 

```{r}
x <- 2:4
y <- 12:14
rbind(x,y)
```

## Factors 

Factors are used to represent categorical data. Factors can be unordered or ordered. One can think of a factor as an integer vector where each integer has a label.

Using factors with labels is better than using integers because factors are self-describing; having a variable that has values 'Male' and 'Female' is better tah having a variable that has values 1 and 2.

```{r}
x <- factor(c('yes', 'yes', 'no', 'no', 'yes'))
x
table(x)
unclass(x)
```

The order of the levels can be set using the levels argument to facotr().

```{r}
x <- factor(c('yes', 'yes', 'no', 'no', 'yes'), levels = c('yes', 'no'))
x
table(x)
unclass(x)

```

Lo que hice acá fue cambiar el valor categórico que tiene las observaciones. yes es 1 en vez de ser 2. 

## Msiing Values 

Missing values are denoted by NA or NaN for undefined mathmatical operations 

is.na() is used to test objects if they are NA 

is.nan() is ued to test for NaN 

NA values have a class also.

A NaN value is also NA but the converse is not true 

```{r}
x <- c(1,2,NA,10,3)
is.na(x)
is.nan(x)

```

```{r}
x<- c(1,2, NaN, NA, 4)
is.na(x)
is.nan(x)

```

## Data Frames 

Special type of list where very element of the list has to have the same length 

They can also store different types of objects 

```{r}
data.matrix(m)

```


```{r}
d <- data.frame(foo = 1:4, bar = c(T,T,F,F))
d
```

